---
title: "Exploration of World Development Indicators"
author: "EHazelton"
date: "July 15, 2018"
output: html_document
---

#note - ideas and inspriation from 
#https://www.kaggle.com/captcalculator/world-development-indicators-sql-dbplyr-highcharts

```{r setup, include=FALSE}
library(tidyverse)

```

```{r import}
# import World development indicators provided by World Bank
# https://datacatalog.worldbank.org/dataset/world-development-indicators
#country_series <- read_csv("wDICountry-Series.csv")
country <- read_csv("WDICountry.csv")
data <- read_csv("WDIData.csv")
#footnote <- read_csv("WDIFootNote.csv")
#series_time <- read_csv("WDISeries-Time.csv")
#series <- read_csv("WDISeries.csv")
```

```{r initial exploring}
# How many unique indicators exist, and what are they? 
indicators <- unique(data$`Indicator Name`)
#write.csv(indicators, "output_indicator_list.csv")

# How are countries grouped by region?
(regions <- unique(country$Region))

# Tidy data
###### Need to add-> spread the indicators; use indicator codes; before, save a indicator name / indicator code cheat sheet
data <- data %>%
  select(-`Country Code`) %>%
  gather(`1960`:`2017`, key = "year", value = "value", convert = T)

# How much USA data do we have?
usa <- data %>%
  filter(`Country Name` == 'United States')
```

Noticed several Co2 emmisions indicators -- what does the general one look like
for the United States?
```{r co2 in USA}
emissions <- filter(usa, `Indicator Name` == 'CO2 emissions (kt)')

# C02 Emissions for USA
ggplot(emissions, aes(year, value)) + 
  geom_line() + 
  labs(title = "CO2 Emissions in United States",
       x = "Year",
       y ="Emissions (kt)")
```

Why is there a dip in the early 1980s in CO2 emissions? And in late 2000s? 

What percentages of Americans do not have electricity? 
```{r hiv}
hiv <- usa %>%
  filter(`Indicator Name` == 'Adults (ages 15+) newly infected with HIV')

#### Update - x ticks to be rounded years
ggplot(hiv, aes(year, value)) +
  geom_point() +
  xlim(c(2007,2017)) +
  labs(title = "US Cases of HIV - newly infected adults (ages 15+)")

```

Although the adult demographic is the only one with numbers, it is exciting
to see that the number of HIV cases is steadily dropping year over year. 

I wonder why we don't have data since 2014? 

# Metric for female labor participation rate, USA? 

# Female education vs male education in USA?

# How does USA female participation rate compare to FPR of other countries?

# How does US armed forces personnel, percent of total, compare to other countries?
### > use the boxy graph thing (where labels are sized according to portion)
# What about percent spending on weapons?
# What about percent export of weapons? 
